---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "75%"
)
```
# TargetedMSM
<!-- badges: start -->
[![R-CMD-check](https://github.com/herbps10/TargetedMSM/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/herbps10/TargetedMSM/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->
  
Targeted estimation for a general class of Non-parametric Marginal Structural Models (NP-MSMs)

## Installation

You can install the development version of TargetedMSM from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("herbps10/TargetedMSM")
```

## Example
Estimate a linear working model with squared error loss function for summarizing a dose-response curve:
```{r example1}
library(TargetedMSM)

set.seed(10016)
data <- simulate_categorical_dose_response(N = 500)

fit <- categorical_dose_response(
  data,
  c("X1", "X2"),
  "A",
  "Y",
  formula = ~1 + A,
  outcome_type = "continuous",
  learners_trt = c("SL.glm.interaction"),
  learners_outcome = c("SL.glm.interaction"),
  loss = loss_squared_error,
  working_model = function(beta, X) beta[1]$mul(X[, 2]$pow(beta[2]))
)

summary(fit)
```

Predict from the estimated NP-MSM and plot results:
```{r plot}
library(purrr)
library(ggplot2)

pred <- data.frame(A = 1:25)
pred$point <- predict(fit, pred)
pred$lower <- map_dbl(predict(fit, pred, type = "joint"), quantile, 0.025)
pred$upper <- map_dbl(predict(fit, pred, type = "joint"), quantile, 0.975)

ggplot(pred, aes(A, point)) +
  geom_line() +
  geom_line(aes(y = lower), lty = 2) +
  geom_line(aes(y = upper), lty = 2) +
  labs(x = "A", y = "Counterfactual mean", title = "Estimated dose-response curve")
```

Estimating a linear working model with squared error loss function for summarizing Conditional Average Treatment Effects:
```{r example2}
set.seed(10016)
data <- simulate_treatment_effect_modification(N = 500)

fit <- treatment_effect_modification(
  data,
  c("X1", "X2"),
  "A",
  "Y",
  formula = ~1 + X2,
  outcome_type = "continuous",
  learners_trt = c("SL.glm.interaction"),
  learners_outcome = c("SL.glm.interaction"),
  loss = loss_squared_error,
  working_model = working_model_linear
)

summary(fit)
```
