---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# TargetedMSM
<!-- badges: start -->
[![R-CMD-check](https://github.com/herbps10/TargetedMSM/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/herbps10/TargetedMSM/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->
  
Targeted estimation for a general class of Non-parametric Marginal Structural Models (NP-MSMs)

## Installation

You can install the development version of TargetedMSM from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("herbps10/TargetedMSM")
```

## Example
Estimating a linear working model with squared error loss function for summarizing Conditional Average Treatment Effects:
```{r example}
library(TargetedMSM)

data <- simulate_treatment_effect_modification(N = 1000, seed = 1)

fit <- treatment_effect_modification(
  data,
  c("X1", "X2"),
  "A",
  "Y",
  formula = ~1 + X2,
  outcome_type = "continuous",
  learners_trt = c("SL.glm.interaction"),
  learners_outcome = c("SL.glm.interaction"),
  working_model = working_model_linear,
  loss = loss_squared_error
)
```

Estimated NP-MSM coefficients:
```{r}
summary(fit)
```
