---
title: "Conditional Average Treatment Effect"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Conditional Average Treatment Effect}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(automsm)
library(ggplot2)
```

This vignette illustrates how to use the `automsm` package to estimate a 
Non-parametric Marginal Structural Model (NP-MSM) to summarize the relationship 
between a Conditional Average Treatment Effect (CATE) and a set of treatment effect modifiers.

The key idea throughout is that the working model (e.g. linear, B-spline, etc.) is
**not** an assumption about the true functional form of the CATE. Instead,
the NP-MSM targets the *best approximation* of the CATE within the working model,
as defined by the loss function. This distinguishes NP-MSMs from classical MSMs:
we make no parametric assumption on the CATE itself, and the 
resulting coefficients are interpreted as the best summary of the CATE
achievable within the chosen working model.

## Two frequentist estimators

The `automsm` package provides two frequentist estimators of NP-MSM parameters: a **one-step** estimator and a **Targeted Minimum Loss-Based Estimator (TMLE)**.
Both estimators are asymptotically normal and efficient, meaning they achieve
the smallest asymptotic variance possible among regular estimators. Both estimators are also double-robust: they remain consistent if either the propensity model or the outcome model is correctly specified.

##  Simulated data
The package has a built in function to generate simulated data for this setting. 
To begin, let's generate $500$ observations.
```{r}
data <- simulate_cate(N = 500, binary = FALSE, seed = 10016)

head(data)
```

The columns of the simulated dataset are

- `X1`, `X2`: continuous covariates.
- `A`: binary treatment variable ($A = 0$ indicates control, and $A = 1$ indicates treatment).
- `Y`: a continuous outcome variable.

For convenience, we will write $X = (X_1, X_2)$ to denote the set of both covariates.

```{r}
ggplot(data, aes(X2, Y, color = factor(A))) +
  geom_point()
```

The _Conditional Average Treatment Effect_ (CATE) is defined as
$\psi(X) = \mathbb{E}[Y \mid A = 1, X] - \mathbb{E}[Y \mid A = 0, X]$. 
Under appropriate causal assumptions, the CATE can be interpreted as the average
causal effect of treatment versus control for those with covariates values $X$. 

For high-dimensional covariates $X$, the CATE may be difficult to estimate. 
One popular parameter for summarizing the CATE is to estimate the _Average Treatment Effect_ 
(ATE), which is the CATE averaged over the population distribution. Formally, the ATE is 
defined as $\psi^{\mathsf{ATE}} = \mathbb{E}[\mathbb{E}[Y \mid A = 1, X] - \mathbb{E}[Y \mid A = 0, X]]$.
In the NP-MSM framework, the ATE corresponds to the special case of a linear working model
containing only an intercept term (that is, when no effect modifiers are included).


## Estimating the ATE

The ATE can be estimated using the `cate` function.
First, we need to set the SuperLearner methods that will be used to estimate the required nuisance parameters.
```{r}
nuisance <- nuisance_control(
  learners_trt = "SL.gam",
  learners_outcome = "SL.gam"
)
```

Then, we can estimate the ATE:
```{r}
fit_ate <- cate(
  data,
  c("X1", "X2"),
  "A",
  "Y",
  formula = ~1,
  outcome_type = "continuous",
  nuisance = nuisance
)
```

The results can be viewed using the `summary` function:
```{r}
summary(fit_ate)
```
As shown in the summary above, the TMLE estimate of the ATE is approximately
`r round(fit_ate$tmle$est[1], 2)` (95% CI: 
`r round(fit_ate$tmle$lower[1], 2)` -- `r round(fit_ate$tmle$upper[1], 2)`).

## Summarizing effect modification
We may also be interested in summarizing how the CATE changes as a functio of the covariates. Suppose we are interested in $X_2$ as a _treatment effect modifier_. 
One way to investigate this is to summarize the relationship between the CATE and $X_2$ in terms of a linear working model that minimizes the squared error between the CATE and  the working model predictions:
```{r}
fit <- cate(
  data,
  c("X1", "X2"),
  "A",
  "Y",
  formula = ~1 + X2,
  outcome_type = "continuous",
  working_model = working_model_linear,
  loss = loss_squared_error,
  nuisance = nuisance
)
```

We can view the estimated NP-MSM coefficients using `summary`:

```{r}
summary(fit)
```

Next, let's use the estimated NP-MSM to make predictions on the treatment effect for a range of values of $X_2$:
```{r}
pred <- data.frame(X2 = seq(0, 1, 0.1))
pred$point <- predict(fit, newdata = pred, estimator = "tmle")

ggplot(pred, aes(X2, point, color = "TMLE Estimate")) +
  geom_line() +
  geom_function(fun = function(x) sin(2 * pi * x), aes(color = "True")) +
  ylab("Treatment Effect") +
  theme(legend.position = "bottom")
```

Because the true CATE is nonlinear, it cannot be represented exactly by a straight line. The linear working model therefore 
does not recover the true CATE; instead, it returns the best linear _summary_ of the CATE under a squared-error loss.
This illustrates a central feature of NP-MSMs: the linear coefficient is interpetable as the best linear projection of the CATE,
and it makes no assumption that the CATE is truly linear. 


We can also make 95\% confidence intervals. Setting `type = "joint"` returns, for 
each row of the new data, a vector of draws from the joint sampling distribution 
of the working-model predictions; we summarize these draws by their 2.5% and
97.5% quantiles to form pointwise 95% intervals.
```{r}
pred$ci <- predict(fit, pred, type = "rvar")
pred$lower <- purrr::map_dbl(pred$ci, quantile, 0.025)
pred$upper <- purrr::map_dbl(pred$ci, quantile, 0.975)

ggplot(pred, aes(X2, point, color = "TMLE Estimate")) +
  geom_line() +
  geom_line(aes(y = lower), lty = 2) +
  geom_line(aes(y = upper), lty = 2) +
  geom_function(fun = function(x) sin(2 * pi * x), aes(color = "True")) +
  ylab("Treatment Effect") +
  labs(color = NULL) +
  theme(legend.position = "bottom")
```

## Comparing the one-step and TMLE estimators

Both frequentist estimators are available from the same fitted object. We can extract each estimator's predictions by passing the `estimator` argument to `predict`, and overlay them to compare. Because both estimators are asymptotically efficient, we expect them to reach substantially the same conclusion, particularly as the sample size grows.

```{r}
compare <- data.frame(X2 = seq(0, 1, 0.1))
compare$tmle <- predict(fit, compare, estimator = "tmle")
compare$onestep <- predict(fit, compare, estimator = "onestep")

ggplot(compare, aes(X2)) +
  geom_line(aes(y = tmle, color = "TMLE")) +
  geom_line(aes(y = onestep, color = "One-step")) +
  geom_function(fun = function(x) sin(2 * pi * x), aes(color = "True")) +
  ylab("Treatment Effect") +
  labs(color = NULL) +
  theme(legend.position = "bottom")
```

We can als ocompare the estimated working-model coefficients directly. The `tidy` method returns a tibble with one row per estimator and working model term, allowing for straightforward comparisons:

```{r}
tidy(fit)
```

The one-step and TMLE coefficient estimates and their confidence intervals should be similar, reflecting that they both share good asymptotic properties. Any differences observed here are finite-sample effects.

## A more flexible working model

The linear summary above doesn't capture the interesting non-linear structure of the treatment effect.
If we wish to capture how the CATE varies with $X_2$ more faithfully, we can project onto
a more flexible working model. A B-spline working model has the same interpretation
as the best approximation of the CATE under a working model for the squared error loss,
but has greater flexibility to capture the nonlinear shape of the TRUE CATE.
```{r}
fit_splines <- cate(
  data,
  c("X1", "X2"),
  "A",
  "Y",
  formula = ~1 + splines::bs(X2, Boundary.knots = c(0, 1), knots = seq(0.25, 0.75, 0.25)),
  outcome_type = "continuous",
  working_model = working_model_linear,
  loss = loss_squared_error,
  nuisance = nuisance
)
```

And now plot the resulting predictions:
```{r}
pred_splines <- data.frame(X2 = seq(0, 1, 0.01))
pred_splines$point <- predict(fit_splines, pred_splines)
pred_splines$ci <- predict(fit_splines, pred_splines, type = "rvar")
pred_splines$lower <- purrr::map_dbl(pred_splines$ci, quantile, 0.025)
pred_splines$upper <- purrr::map_dbl(pred_splines$ci, quantile, 0.975)

ggplot(pred_splines, aes(X2, point, color = "TMLE Estimate")) +
  geom_line() +
  geom_line(aes(y = lower), lty = 2) +
  geom_line(aes(y = upper), lty = 2) +
  geom_function(fun = function(x) sin(2 * pi * x), aes(color = "True")) +
  ylab("Treatment Effect") +
  labs(color = NULL) +
  theme(legend.position = "none")
```

Unlike in the linear summary, the B-spline fit recovers the rising-and-falling shape
of the true CATE, demonstrating the advantage of a richer working model.
